---
title: "Nighttime Lights Analysis"
author: "Data Lab"
date: "9/20/2024"
format:
  html:
    toc: true
    code-fold: true
knitr:
  opts_chunk:
    warning: false
    message: false
---

Nighttime lights have become a commonly used resource to estimate changes in local economic activity. This document analyzes nighttime lights within the country.

# Data

We use nighttime lights data from [VIIRS Black Marble](https://blackmarble.gsfc.nasa.gov/). Raw nighttime lights data requires correction due to cloud cover and stray light, such as lunar light. The Black Marble dataset applies [advanced algorithms](https://www.sciencedirect.com/science/article/pii/S003442571830110X) to correct raw nighttime light values and calibrate data so that trends in lights over time can be meaningfully analyzed. From VIIRS Black Marble, we use data from January 2012 through present—where data is available at a 500-meter resolution.

# Methodology

Within different units of analysis (e.g, administrative units) we use the sum of nighttime lights. Where relevant, we distinguish between lights generated from gas flaring and lights removing gas flaring. We use the [World Bank's Global Gas Flaring Tracker](https://www.worldbank.org/en/programs/gasflaringreduction/global-flaring-data) which indicates the location of gas flaring locations. When removing gas flaring lights, we remove lights within 10km of a gas flaring location; when looking at lights in gas flaring locations, we take the sum of lights within 10km of gas flaring locations. 

# Code Setup

```{r}
#| warning: false
#| message: false

#### MAIN PARAMS
proj_dir     <- "~/Dropbox/World Bank/Side Work/Syria Economic Monitor 2024"
iso3_code    <- "SYR"
iso2_code    <- "SY"
pc_base_year <- 2019

DAILY_START <- "2023-01-01"

#### Packages
library(tidyverse)
library(sf)
library(leaflet)
library(leaflet.providers)
library(ggpubr)
library(terra)
library(tidyterra)
library(gtools)
library(readxl)
library(janitor)
library(geodata)
library(exactextractr)
library(blackmarbler)
library(dplyr)
library(readxl)
library(janitor)
library(ggpubr)
library(WDI)
library(kableExtra)
library(elevatr)
library(ggnewscale)
library(pals)
library(haven)
library(raster)

#### Paths

## Define
data_dir       <- file.path(proj_dir, "data")
ntl_bm_dir     <- file.path(data_dir, "Nighttime Lights BlackMarble")
gadm_dir       <- file.path(data_dir, "GADM")
gas_flare_dir  <- file.path(data_dir, "gas-flaring")

## Create
dir.create(proj_dir)
dir.create(data_dir)
dir.create(ntl_bm_dir)
dir.create(gadm_dir)
dir.create(gas_flare_dir)

#### Bearer
nasa_bearer <- read_csv("~/Dropbox/bearer_bm.csv") %>%
  pull(token)

#### Theme
theme_manual <- theme_classic2() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))

#### GADM Boundaries
# for(i in 0:3){
#   for(iso3_i in iso3_code){
#     gadm(iso3_code, level=i, path = file.path(gadm_dir))
#   }
# } 

#### Load GADM Functions
load_gadm <- function(i){
  roi_sf <- file.path(gadm_dir, 'gadm') %>%
    list.files(full.names = T) %>%
    str_subset(paste0("_",i,"_pk.rds")) %>%
    map_df(function(file_path){
      readRDS(file_path) %>%
        st_as_sf()
    })
}


#### Load Boundaries
load_boundaries <- function(i){
  if(i == 0){
    roi_sf <- st_read(file.path(data_dir, "Syria Boundaries", "syr_admbnda_adm0_uncs_unocha_20201217.shp")) %>%
      mutate(uid = 1:n())
  }
  
  if(i == 1){
    roi_sf <- st_read(file.path(data_dir, "Syria Boundaries", "syr_admbnda_adm1_uncs_unocha_20201217.shp")) %>%
      mutate(uid = 1:n())
  }
  
  if(i == 2){
    roi_sf <- st_read(file.path(data_dir, "Syria Boundaries", "syr_admbnda_adm2_uncs_unocha_20201217.shp")) %>%
      mutate(uid = 1:n())
  }
  
  if(i == 3){
    roi_sf <- st_read(file.path(data_dir, "Syria Boundaries", "syr_admbnda_adm3_uncs_unocha_20201217.shp")) %>%
      mutate(uid = 1:n())
  }
  
  if(i == 4){
    roi_sf <- st_read(file.path(data_dir, "Syria Boundaries", "syria_adm4_5km.geojson")) %>%
      mutate(uid = 1:n()) %>%
      dplyr::select(uid, admin4Name_en)
  }
  
  return(roi_sf)
}

#### Elevation
dir.create(file.path(data_dir, "elevation"))

OUT_PATH <- file.path(data_dir, "elevation", "elev.tiff")

if(!file.exists(OUT_PATH)){
  
  if(length(iso3_code) > 1){
    
    # elev_list <- lapply(iso3_code, function(iso3_code_i){
    #   elevation_3s(country=iso3_code_i, path=tempdir())
    # }) 
    # elev_r <- do.call(mosaic, c(elev_list, fun="mean"))
    
    elev_list <- lapply(iso3_code, function(iso3_code_i){
      
      adm0_sf <- gadm(iso3_code, level=0, path = tempdir())
      get_elev_raster(locations = adm0_sf, z = 8)
      
    })
    elev_r <- do.call(mosaic, c(elev_list, fun="mean"))
    
  } else{
    adm0_sf <- load_gadm(0)
    
    #elev_r <- elevation_3s(country=iso3_code, path=tempdir())
    elev_r <- get_elev_raster(locations = adm0_sf, z = 8)
    
  }
  
  writeRaster(elev_r, OUT_PATH, overwrite = T)
  
}

#### NTL Months to Download
ntl_months <- seq.Date(ymd("2012-01-01"), to = ymd("2024-11-01"), by = "month") %>%
  as.character()

ntl_months <- seq.Date(ymd("2018-01-01"), to = ymd("2024-11-01"), by = "month") %>%
  as.character()

ntl_days <- seq.Date(from = ymd("2024-01-01"), to = ymd("2024-01-03"), by = "day") %>%
  as.character()

#### Colors
wbg_color_light <- "#1AA1DB"
wbg_color_dark <- "#02224B"
```

# Prep Data

## Download Cities

We use a dataset from [Geonames](https://public.opendatasoft.com/explore/dataset/geonames-all-cities-with-a-population-1000/export/?flg=en-us&disjunctive.cou_name_en&sort=name) which indicates the location of cities with over 1,000 people. 

```{r}
#| warning: false
#| message: false

dir.create(file.path(data_dir, "cities"))

## Cities
OUT_FILE_ALL <- file.path(data_dir, "cities", "cities_all.Rds")

if(!file.exists(OUT_FILE_ALL)){
  # https://public.opendatasoft.com/explore/dataset/geonames-all-cities-with-a-population-1000/export/?flg=en-us&disjunctive.cou_name_en&sort=name
  city_df <- read_delim("https://public.opendatasoft.com/api/explore/v2.1/catalog/datasets/geonames-all-cities-with-a-population-1000/exports/csv?lang=en&timezone=America%2FNew_York&use_labels=true&delimiter=%3B", delim = ";")
  
  saveRDS(city_df, OUT_FILE_ALL)
} 

## Country Cities
OUT_FILE_COUNTRY_DF      <- file.path(data_dir, "cities", "cities_country_df.Rds")
OUT_FILE_COUNTRY_SF      <- file.path(data_dir, "cities", "cities_country_sf.Rds")
OUT_FILE_COUNTRY_5K_SF   <- file.path(data_dir, "cities", "cities_country_5km_sf.Rds")
OUT_FILE_COUNTRY_10K_SF  <- file.path(data_dir, "cities", "cities_country_10km_sf.Rds")
OUT_FILE_COUNTRY_BUFF_SF <- file.path(data_dir, "cities", "cities_country_buff_sf.Rds")

if(!file.exists(OUT_FILE_COUNTRY_SF)){
  
  ## Cleanup
  city_country_df <- city_df[city_df$`Country Code` %in% iso2_code,] %>%
    clean_names()
  
  city_country_df <- city_country_df %>%
    separate(coordinates, into = c("lat", "lon"), sep = ",")
  
  ## Spatially Define
  city_country_sf <- city_country_df %>%
    st_as_sf(coords = c("lon", "lat"),
             crs = 4326)
  
  ## Buffer
  city_country_5km_sf <- city_country_sf %>%
    st_buffer(dist = 5000)
  
  city_country_10km_sf <- city_country_sf %>%
    st_buffer(dist = 10000)
  
  ## Dynamic Buffer
  city_country_sf <- city_country_sf %>%
    dplyr::mutate(buffer = case_when(
      population >= 10000000 ~ 30,
      population >= 5000000 ~ 25,
      population >= 1000000 ~ 20,
      population >= 500000 ~ 15,
      population >= 100000 ~ 10,
      TRUE ~ 5
    ))
  
  city_sub_30km_buff_sf <- city_country_sf[city_country_sf$buffer %in% 30,] %>% st_buffer(dist = 30*1000)
  city_sub_25km_buff_sf <- city_country_sf[city_country_sf$buffer %in% 25,] %>% st_buffer(dist = 25*1000)
  city_sub_20km_buff_sf <- city_country_sf[city_country_sf$buffer %in% 20,] %>% st_buffer(dist = 20*1000)
  city_sub_15km_buff_sf <- city_country_sf[city_country_sf$buffer %in% 15,] %>% st_buffer(dist = 15*1000)
  city_sub_10km_buff_sf <- city_country_sf[city_country_sf$buffer %in% 10,] %>% st_buffer(dist = 10*1000)
  city_sub_5km_buff_sf  <- city_country_sf[city_country_sf$buffer %in% 5,]  %>% st_buffer(dist = 5*1000)
  
  city_buff_sf <- bind_rows(city_sub_30km_buff_sf,
                            city_sub_25km_buff_sf,
                            city_sub_20km_buff_sf,
                            city_sub_15km_buff_sf,
                            city_sub_10km_buff_sf,
                            city_sub_5km_buff_sf)
  
  
  ## Export
  saveRDS(city_country_df, OUT_FILE_COUNTRY_DF)
  
  saveRDS(city_country_5km_sf,  OUT_FILE_COUNTRY_5K_SF)
  saveRDS(city_country_10km_sf, OUT_FILE_COUNTRY_10K_SF)
  saveRDS(city_buff_sf,         OUT_FILE_COUNTRY_BUFF_SF)
  saveRDS(city_country_sf,      OUT_FILE_COUNTRY_SF)
  
} 

```


## Download and Append gas flaring

We download and append gas flaring data from the [World Bank Global Gas Flaring Database](https://datacatalog.worldbank.org/search/dataset/0037743).

```{r}
#| warning: false
#| message: false

dir.create(file.path(gas_flare_dir, "rawdata"))
dir.create(file.path(gas_flare_dir, "finaldata"))

#### Download data
# https://datacatalog.worldbank.org/search/dataset/0037743
if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0045623/viirs_global_flaring_d.7_slope_0.029353_2017_web_v1.xlsx?versionId=2023-01-18T20:03:32.2273754Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0045622/viirs_global_flaring_d.7_slope_0.029353_2018_web.xlsx?versionId=2023-01-18T20:02:43.3965005Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2019.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0045621/viirs_global_flaring_d.7_slope_0.029353_2019_web_v20201114-3.xlsx?versionId=2023-01-18T20:03:09.2456111Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2019.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0084248/2020%20Global%20Gas%20Flaring%20Volumes.xlsx?versionId=2023-01-18T20:03:53.8309309Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2021.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0087112/2021%20Global%20Gas%20Flaring%20Volumes.xlsx?versionId=2023-01-18T20:02:21.4951166Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2021.xlsx"), 
                mode = "wb")
}

#### Append data
clean_data <- function(x){
  x %>% 
    clean_names() %>% 
    dplyr::filter(iso_code %in% iso3_code)
} 

df_2021 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2021.xlsx"), 2) %>% clean_data()

df_2020_1 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 1) %>% clean_data()
df_2020_2 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 2) %>% clean_data()
df_2020_3 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 3) %>% clean_data()

df_2019 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2019.xlsx"), 1) %>% clean_data()

df_2018_4 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 4) %>% clean_data()
df_2018_5 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 5) %>% clean_data()
df_2018_6 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 6) %>% clean_data()

df_2017_1 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 1) %>% clean_data()
df_2017_2 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 2) %>% clean_data()
df_2017_3 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 3) %>% clean_data()

gs_df <- bind_rows(
  df_2021,
  df_2020_1,
  df_2020_2,
  df_2020_3,
  df_2019,
  df_2018_4,
  df_2018_5,
  df_2018_6,
  df_2017_1,
  df_2017_2,
  df_2017_3
)

gs_df <- gs_df %>%
  dplyr::select(latitude, longitude) %>%
  distinct() %>%
  dplyr::mutate(uid = 1:n())

gs_sf <- gs_df %>% 
  st_as_sf(coords = c("longitude", "latitude"),
           crs = 4326)

gs_5km_sf <- gs_sf %>%
  st_buffer(dist = 5000)

gs_10km_sf <- gs_sf %>%
  st_buffer(dist = 10000)

saveRDS(gs_df, file.path(gas_flare_dir, "finaldata", "gas_flare_locations.Rds"))
saveRDS(gs_5km_sf, file.path(gas_flare_dir, "finaldata", "gas_flare_locations_5km.Rds"))
saveRDS(gs_10km_sf, file.path(gas_flare_dir, "finaldata", "gas_flare_locations_10km.Rds"))
```

## Prep ADM4 File

```{r}
if(F){
  roi_sf <- read_sf(file.path(data_dir, 
                              "Syria Boundaries",
                              "syr_admin_20240711.gdb"),
                    "syr_pplp_humanitarian_location_unocha")
  roi_sf <- roi_sf %>%
    st_buffer(5000)
  
  write_sf(roi_sf, file.path(data_dir, 
                             "Syria Boundaries",
                             "syria_adm4_5km.geojson"),
           delete_dsn = T,
           delete_layer = T)
}

```


## Download Nighttime Lights

```{r}
#| warning: false
#| message: false

#### NTL
roi_sf <- load_gadm(0)

dir.create(file.path(ntl_bm_dir, "rasters"))
dir.create(file.path(ntl_bm_dir, "rasters", "annual"))
dir.create(file.path(ntl_bm_dir, "rasters", "monthly"))
dir.create(file.path(ntl_bm_dir, "rasters", "daily"))

r <- bm_raster(roi_sf = roi_sf,
               product_id = "VNP46A4",
               date = 2012:2023,
               bearer = nasa_bearer,
               output_location_type = "file",
               file_dir = file.path(ntl_bm_dir, "rasters", "annual"))

r <- bm_raster(roi_sf = roi_sf,
               product_id = "VNP46A3",
               date = ntl_months,
               bearer = nasa_bearer,
               output_location_type = "file",
               file_dir = file.path(ntl_bm_dir, "rasters", "monthly"))

# r <- bm_raster(roi_sf = roi_sf,
#                product_id = "VNP46A2",
#                date = ntl_days,
#                bearer = nasa_bearer,
#                output_location_type = "file",
#                file_dir = file.path(ntl_bm_dir, "rasters", "daily"))

```

## Aggregate Nighttime Lights
```{r}
#| warning: false
#| message: false

# Individual Files -------------------------------------------------------------
dir.create(file.path(ntl_bm_dir, "aggregated_individual_files"))

#### Setup loops
for(unit in c("adm0", 
              "adm1", 
              "adm2",
              "adm3",
              "adm4",
              "adm4_5km",
              "bordercrossings",
              "cities5km",
              "cities10km",
              "citiesbuff")){
  for(time_type in c("annual", "monthly", "daily")){
    
    if(time_type == "annual"){
      time_vec <- 2012:2023
    }
    
    if(time_type == "monthly"){
      time_vec <- ntl_months
    }
    
    if(time_type == "daily"){
      #time_vec <- ntl_days
      
      time_vec <- file.path(ntl_bm_dir, "rasters", "daily") %>%
        list.files() %>%
        str_replace_all("VNP46A2_Gap_Filled_DNB_BRDF-Corrected_NTL_qflag_t", "") %>%
        str_replace_all(".tif", "") %>%
        str_replace_all("_", "-")
      
    }
    
    for(time_i in time_vec){
      
      #### Only process if file doesn't exist
      dir.create(file.path(ntl_bm_dir, "aggregated_individual_files", unit))
      dir.create(file.path(ntl_bm_dir, "aggregated_individual_files", unit, time_type))
      
      OUT_FILE <- file.path(ntl_bm_dir, "aggregated_individual_files", unit, time_type, 
                            paste0("ntl_", time_i, ".Rds"))
      
      if(!file.exists(OUT_FILE)){
        
        #### Load Unit
        if(unit == "adm0"){
          roi_sf <- load_boundaries(0)
        }
        
        if(unit == "adm1"){
          roi_sf <- load_boundaries(1)
        }
        
        if(unit == "adm2"){
          roi_sf <- load_boundaries(2)
        }
        
        if(unit == "adm3"){
          roi_sf <- load_boundaries(3)
        }
        
        if(unit == "adm4_5km"){
          roi_sf <- load_boundaries(4)

        }
        
        if(unit == "bordercrossings"){
          roi_sf <- read_csv(file.path(data_dir, "Border Crossing Locations - All", 
                                       "RawData",
                                       "syria_border_crossings.csv")) %>%
            st_as_sf(coords = c("longitude", "latitude"),
                     crs = 4326) %>%
            st_buffer(1500)
        }
        
        if(unit == "cities5km"){
          roi_sf <- readRDS(file.path(data_dir, "cities", "cities_country_5km_sf.Rds"))
        }
        
        if(unit == "cities10km"){
          roi_sf <- readRDS(file.path(data_dir, "cities", "cities_country_10km_sf.Rds"))
        }
        
        if(unit == "citiesbuff"){
          roi_sf <- readRDS(file.path(data_dir, "cities", "cities_country_buff_sf.Rds"))
        }
        
        
        #### Load Raster
        if(time_type == "annual"){
          bm_product <- "VNP46A4"
          time_raster_i <- time_i 
        }
        
        if(time_type == "monthly"){
          bm_product <- "VNP46A3"
          
          time_raster_i <- time_i %>% 
            substring(1, 7) %>%
            str_replace_all("-", "_")
        }
        
        if(time_type == "daily"){
          bm_product <- "VNP46A2"
          
          time_raster_i <- time_i %>% 
            substring(1, 10) %>%
            str_replace_all("-", "_")
        }
        
        if(time_type %in% c("daily")){
          ntl_r <- rast(file.path(ntl_bm_dir, "rasters", time_type, 
                                  paste0("VNP46A2_Gap_Filled_DNB_BRDF-Corrected_NTL_qflag_t",time_raster_i,".tif")))
        } else{
          ntl_r <- rast(file.path(ntl_bm_dir, "rasters", time_type, 
                                  paste0(bm_product,"_NearNadir_Composite_Snow_Free_qflag_t",time_raster_i,".tif")))
        }
        
        #### Gas Flaring Mask
        gs_5km_sf  <- readRDS(file.path(gas_flare_dir, "finaldata", "gas_flare_locations_5km.Rds"))
        gs_10km_sf <- readRDS(file.path(gas_flare_dir, "finaldata", "gas_flare_locations_10km.Rds"))
        
        ntl_gf_5km_r   <- mask(ntl_r, gs_5km_sf)
        ntl_nogf_5km_r <- mask(ntl_r, gs_5km_sf, inverse = T)
        
        ntl_gf_10km_r   <- mask(ntl_r, gs_10km_sf)
        ntl_nogf_10km_r <- mask(ntl_r, gs_10km_sf, inverse = T)
        
        #### Extract NTL
        roi_sf$ntl_sum          <- exact_extract(ntl_r, roi_sf, "sum")
        
        roi_sf$ntl_gf_5km_sum   <- exact_extract(ntl_gf_5km_r, roi_sf, "sum")
        roi_sf$ntl_nogf_5km_sum <- exact_extract(ntl_nogf_5km_r, roi_sf, "sum")
        
        roi_sf$ntl_gf_10km_sum   <- exact_extract(ntl_gf_10km_r, roi_sf, "sum")
        roi_sf$ntl_nogf_10km_sum <- exact_extract(ntl_nogf_10km_r, roi_sf, "sum")
        
        #### Cleanup
        roi_sf$date <- time_i
        
        roi_df <- roi_sf %>%
          st_drop_geometry()
        
        #### Export
        saveRDS(roi_df, OUT_FILE)
        
      }
    }
  }
}

#### Append Files
dir.create(file.path(ntl_bm_dir, "aggregated"))

for(unit in c("adm0", 
              "adm1", 
              "adm2",
              "adm3",
              "adm4_5km",
              "bordercrossings",
              "cities5km",
              "cities10km",
              "citiesbuff")){
  for(time_type in c("annual", "monthly", "daily")){
    
    ntl_df <- file.path(ntl_bm_dir, "aggregated_individual_files", unit, time_type) %>%
      list.files(full.names = T,
                 pattern = ".Rds") %>%
      map_df(readRDS)
    
    if(unit == "bordercrossings"){
      ntl_df <- ntl_df %>%
        dplyr::select(-c("...6", "...7",
                         "...8", "...9"))
    }
    
    saveRDS(ntl_df,   file.path(ntl_bm_dir, "aggregated", paste0("ntl_", unit, "_", time_type, ".Rds")))
    write_csv(ntl_df, file.path(ntl_bm_dir, "aggregated", paste0("ntl_", unit, "_", time_type, ".csv")))
    write_dta(ntl_df, file.path(ntl_bm_dir, "aggregated", paste0("ntl_", unit, "_", time_type, ".dta")))
    
  }
}

```

# Maps of Nighttime Lights

## Interactive Map

```{r}

#| warning: false
#| message: false

## Load boundaries
adm0_sf <- load_gadm(0)

## Load/prep raster
prep_r <- function(year_i){
  r <- rast(file.path(ntl_bm_dir, "rasters", "annual",
                      paste0("VNP46A4_NearNadir_Composite_Snow_Free_qflag_t",year_i,".tif")))
  r <- r %>% mask(adm0_sf)
  r[][r[] <= 0.05] <- NA
  r[] <- log(r[] + 1)
  #r[] <- log(r[] + 1)
  return(r)
}

r_2012 <- prep_r(2012)
r_2013 <- prep_r(2013)
r_2014 <- prep_r(2014)
r_2015 <- prep_r(2015)
r_2016 <- prep_r(2016)
r_2017 <- prep_r(2017)
r_2018 <- prep_r(2018)
r_2019 <- prep_r(2019)
r_2020 <- prep_r(2020)
r_2021 <- prep_r(2021)
r_2022 <- prep_r(2022)
r_2023 <- prep_r(2023)

## Make map
pal <- colorNumeric(c("black", "yellow", "red"), unique(c(r_2012[],
                                                          r_2013[],
                                                          r_2014[],
                                                          r_2015[],
                                                          r_2016[],
                                                          r_2017[],
                                                          r_2018[],
                                                          r_2019[],
                                                          r_2020[],
                                                          r_2021[],
                                                          r_2022[],
                                                          r_2023[])),
                    na.color = "transparent")

leaflet() %>%
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addRasterImage(r_2012, colors = pal, opacity = 1, group = "2012") %>%
  addRasterImage(r_2013, colors = pal, opacity = 1, group = "2013") %>%
  addRasterImage(r_2014, colors = pal, opacity = 1, group = "2014") %>%
  addRasterImage(r_2015, colors = pal, opacity = 1, group = "2015") %>%
  addRasterImage(r_2016, colors = pal, opacity = 1, group = "2016") %>%
  addRasterImage(r_2017, colors = pal, opacity = 1, group = "2017") %>%
  addRasterImage(r_2018, colors = pal, opacity = 1, group = "2018") %>%
  addRasterImage(r_2019, colors = pal, opacity = 1, group = "2019") %>%
  addRasterImage(r_2020, colors = pal, opacity = 1, group = "2020") %>%
  addRasterImage(r_2021, colors = pal, opacity = 1, group = "2021") %>%
  addRasterImage(r_2022, colors = pal, opacity = 1, group = "2022") %>%
  addRasterImage(r_2023, colors = pal, opacity = 1, group = "2023") %>%
  addLayersControl(
    baseGroups = paste0(2012:2023),
    options = layersControlOptions(collapsed=FALSE)
  )
```


## Static Map of Nighttime Lights

```{r}
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

## Load boundaries
adm0_sf <- load_gadm(0)

## Elevation
elev_r <- rast(file.path(data_dir, "elevation", "elev.tiff"))
slope_r <- terrain(elev_r)
slope_r <- slope_r %>% mask(adm0_sf)

## Latest Year
r <- rast(file.path(ntl_bm_dir, "rasters", "annual",
                    paste0("VNP46A4_NearNadir_Composite_Snow_Free_qflag_t",2023,".tif")))
r <- r %>% mask(adm0_sf)

r[][r[] <= 1] <- NA
r[] <- log(r[] + 1)
r[] <- log(r[] + 1)

ocean_colors <- ocean.deep(100)
ocean_colors[98:100] <- "#FFFFFF"

ggplot() +
  geom_spatraster(data = slope_r, 
                  show.legend = FALSE,
                  maxcell = 9e+07) +
  scale_fill_gradient(low = "#49232E",
                      high = "black",
                      na.value = "transparent") +
  ggnewscale::new_scale_fill() +
  geom_spatraster(data = r,
                  aes(fill = t2023),
                  maxcell = 9e+07) +
  scale_fill_gradientn(colors = ocean_colors,
                       #midpoint = 2,
                       na.value = "transparent") +
  labs(fill = "Light Intensity",
       title = "Nighttime Lights",
       subtitle = 2023) +
  coord_sf() +
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
        plot.subtitle = element_text(face = "italic", hjust = 0.5, size = 16),
        legend.position = c(0.8, 0.25)) +
  guides(fill = guide_colorbar(direction = "horizontal",
                               label = FALSE,         
                               title.position = "top"))
```
# Diagnostics

## Nighttime Lights: Gas Flaring vs Non Gas Flaring

### Trend Figure
```{r}
#| warning: false
#| message: false

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm0_annual.Rds"))

ntl_df %>%
  dplyr::select(date, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -date) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares",
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares"
  )) %>%
  
  ggplot() +
  geom_col(aes(x = date,
               y = value,
               fill = name),
           stat = "identity", position = 'dodge') +
  labs(fill = NULL,
       x = NULL,
       y = "NTL") +
  scale_fill_manual(values = c(wbg_color_light,
                               wbg_color_dark)) +
  theme_manual
```

### Percent of NTL from Gas Flaring
```{r}
#| warning: false
#| message: false

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm0_annual.Rds"))

ntl_df %>%
  dplyr::mutate(per_gf = (ntl_gf_10km_sum / ntl_sum) * 100) %>%
  dplyr::select(date, ADM0_EN, per_gf) %>%
  pivot_wider(id_cols = date,
              names_from = ADM0_EN,
              values_from = per_gf) %>%
  dplyr::rename(Year = date) %>%
  kable()
```

## NTL vs City Population

We use a dataset from [Geonames] that maps the locations of all cities with other 1,000 people. We extract total nighttime lights withing a 5km buffer of each city. The below figure compares nighttime lights with population. 

```{r}
city_df <- readRDS(file.path(data_dir, "cities", "cities_country_df.Rds"))
city_df <- city_df %>%
  dplyr::select(geoname_id, lat, lon)

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "citiesbuff", "_", "annual", ".Rds")))

ntl_coord_df <- ntl_df %>%
  dplyr::filter(date == 2023) %>%
  left_join(city_df, by = "geoname_id") %>%
  dplyr::mutate(lat = lat %>% as.numeric(),
                lon = lon %>% as.numeric()) %>%
  arrange(ntl_sum)

adm0_sf <- load_gadm(0)

ntl_df %>%
  ggplot() +
  geom_sf(data = adm0_sf) +
  geom_point(data = ntl_coord_df,
             aes(x = lon, 
                 y = lat,
                 fill = log(ntl_sum),
                 size = ntl_sum),
             pch = 21,
             color = "black") +
  labs(size = "Nighttime\nLights",
       title = "Nighttime Lights Across Cities") +
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5)) +
  scale_fill_gradient(low = wbg_color_dark,
                      high = wbg_color_light) +
  guides(fill = "none")

```


```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "cities10km", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::filter(population > 0,
                date == 2023) %>%
  ggplot(aes(x = log(ntl_sum),
             y = log(population))) +
  geom_point(fill = wbg_color_dark) +
  geom_smooth(method=lm, 
              se=FALSE,
              color = wbg_color_light) +
  stat_cor() +
  theme_minimal() +
  labs(x = "NTL, Logged",
       y = "Population, Logged")
```


## Nighttime Lights as Proxy for GDP
```{r}
# Download GDP
dir.create(file.path(data_dir, "wdi"))
OUT_FILE <- file.path(data_dir, "wdi", "gdp.Rds")

if(!file.exists(OUT_FILE)){
  gdp_df <- WDI(indicator = c("NY.GDP.PCAP.CD",
                              "NY.GDP.MKTP.CD"), 
                country = iso3_code, 
                start = 2012, 
                end = 2023)
  
  saveRDS(gdp_df, OUT_FILE)
} else{
  gdp_df <- readRDS(OUT_FILE)
}

# Correlation with GDP
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "annual", ".Rds")))

ntl_df <- ntl_df %>%
  dplyr::rename(year = date)

# Filter
if(paste(iso3_code, collapse = ";") %in% "SSD;SDN"){
  ntl_df <- ntl_df[ntl_df$COUNTRY %in% "Sudan",]
  gdp_df <- gdp_df[gdp_df$country %in% "Sudan",]
}

ntl_gdp_df <- ntl_df %>%
  left_join(gdp_df, by = "year") %>%
  clean_names()

## Line Plots
ntl_gdp_df %>%
  dplyr::select(year, 
                ntl_sum,
                ntl_gf_10km_sum,
                ntl_nogf_10km_sum,
                ny_gdp_pcap_cd,
                ny_gdp_mktp_cd) %>%
  pivot_longer(cols = -year) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_sum" ~ "Nighttime Lights",
    name == "ntl_gf_10km_sum" ~ "Nighttime Lights:\nGas Flares",
    name == "ntl_nogf_10km_sum" ~ "Nighttime Lights:\nNo Gas Flares",
    name == "ny_gdp_pcap_cd" ~ "GDP, Per Capita",
    name == "ny_gdp_mktp_cd" ~ "GDP"
  )) %>%
  dplyr::mutate(type = name %>% str_detect("GDP")) %>%
  ggplot(aes(x = year,
             y = value,
             color = type)) +
  geom_line(linewidth = 1) +
  facet_wrap(~name,
             scales = "free_y") +
  labs(x = NULL,
       y = NULL) +
  scale_color_manual(values = c(wbg_color_light, wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "none")

## Correlation: GDP
ntl_gdp_df %>%
  dplyr::select(ntl_sum, 
                ntl_gf_10km_sum,
                ntl_nogf_10km_sum,
                ny_gdp_mktp_cd) %>%
  pivot_longer(cols = -c(ny_gdp_mktp_cd)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_sum" ~ "NTL",
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flare",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flare",
  )) %>%
  
  ggplot(aes(x = value,
             y = ny_gdp_mktp_cd)) +
  geom_point() +
  geom_smooth(method=lm, 
              se=FALSE,
              color = wbg_color_light) +
  stat_cor() +
  labs(x = "NTL",
       y = "GDP",
       title = "Correlation of Nighttime Lights with GDP") +
  facet_wrap(~name,
             scales = "free") +
  theme_manual 

## Correlation: GDP
ntl_gdp_df %>%
  dplyr::select(ntl_sum, 
                ntl_gf_10km_sum,
                ntl_nogf_10km_sum,
                ny_gdp_pcap_cd) %>%
  pivot_longer(cols = -c(ny_gdp_pcap_cd)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_sum" ~ "NTL",
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flare",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flare",
  )) %>%
  
  ggplot(aes(x = value,
             y = ny_gdp_pcap_cd)) +
  geom_point() +
  geom_smooth(method=lm, 
              se=FALSE,
              color = wbg_color_light) +
  stat_cor() +
  labs(x = "NTL",
       y = "GDP, per Capita",
       title = "Correlation of Nighttime Lights with per Capita GDP") +
  facet_wrap(~name,
             scales = "free") +
  theme_manual
```


# Trends in Nighttime Lights

## Country Level

### Annual

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "annual", ".Rds")))

ntl_df %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM0_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::select(date, ADM0_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, ADM0_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM0_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

### Monthly

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "monthly", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd("2018-01-01")) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM0_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "monthly", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd("2018-01-01")) %>%
  dplyr::select(date, ADM0_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, ADM0_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM0_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

### Daily

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "daily", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd(DAILY_START)) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM0_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "daily", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd(DAILY_START)) %>%
  dplyr::select(date, ADM0_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, ADM0_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM0_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

## ADM 1: Top ADM Units

### Annual

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm1", "_", "annual", ".Rds")))

top_adm <- ntl_df %>%
  filter(date == 2023) %>%
  arrange(-ntl_sum) %>%
  pull(uid) %>%
  head(12)

ntl_df %>%
  filter(uid %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM1_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual 
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm1", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::select(date, uid, ADM1_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, uid, ADM1_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  filter(uid %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM1_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

### Monthly

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm1", "_", "monthly", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd("2018-01-01")) %>%
  filter(uid %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM1_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual 
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm1", "_", "monthly", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd("2018-01-01")) %>%
  dplyr::select(date, uid, ADM1_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, uid, ADM1_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  filter(uid %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM1_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

### Daily

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm1", "_", "daily", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd(DAILY_START)) %>%
  filter(uid %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM1_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual 
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm1", "_", "daily", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd(DAILY_START)) %>%
  dplyr::select(date, uid, ADM1_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, uid, ADM1_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  filter(uid %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM1_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

## ADM 2: Top ADM Units

### Annual

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm2", "_", "annual", ".Rds")))

top_adm <- ntl_df %>%
  filter(date == 2023) %>%
  arrange(-ntl_sum) %>%
  pull(ADM2_EN) %>%
  head(12)

ntl_df %>%
  filter(ADM2_EN %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM2_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm2", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::select(date, ADM2_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, ADM2_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  filter(ADM2_EN %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM2_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

### Monthly

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm2", "_", "monthly", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd("2018-01-01")) %>%
  filter(ADM2_EN %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM2_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm2", "_", "monthly", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd("2018-01-01")) %>%
  dplyr::select(date, ADM2_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, ADM2_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  filter(ADM2_EN %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM2_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

### Daily

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm2", "_", "daily", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd(DAILY_START)) %>%
  filter(ADM2_EN %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM2_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm2", "_", "daily", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd(DAILY_START)) %>%
  dplyr::select(date, ADM2_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, ADM2_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  filter(ADM2_EN %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM2_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

## ADM 3: Top ADM Units

### Annual

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm3", "_", "annual", ".Rds")))

top_adm <- ntl_df %>%
  filter(date == 2023) %>%
  arrange(-ntl_sum) %>%
  pull(ADM3_EN) %>%
  head(12)

ntl_df %>%
  filter(ADM3_EN %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM3_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm3", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::select(date, ADM3_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, ADM3_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  filter(ADM3_EN %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM3_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

### Monthly

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm3", "_", "monthly", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd("2018-01-01")) %>%
  filter(ADM3_EN %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM3_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm3", "_", "monthly", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd("2018-01-01")) %>%
  dplyr::select(date, ADM3_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, ADM3_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  filter(ADM3_EN %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM3_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

### Daily

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm3", "_", "daily", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd(DAILY_START)) %>%
  filter(ADM3_EN %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~ADM3_EN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm3", "_", "daily", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd(DAILY_START)) %>%
  dplyr::select(date, ADM3_EN, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, ADM3_EN)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  filter(ADM3_EN %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~ADM3_EN) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

## Border Crossings

### Annual

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "bordercrossings", "_", "annual", ".Rds")))

ntl_df %>%
  filter(border_country %in% "Iraq") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Iraq") +
  theme_manual

ntl_df %>%
  filter(border_country %in% "Jordan") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Jordan") +
  theme_manual

ntl_df %>%
  filter(border_country %in% "Lebanon") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Lebanon") +
  theme_manual

ntl_df %>%
  filter(border_country %in% "Turkey") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Turkey") +
  theme_manual
```

```{r}
# ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
#                             paste0("ntl_", "bordercrossings", "_", "annual", ".Rds")))
# 
# ntl_df %>%
#   dplyr::rename(city_name = name) %>%
#   dplyr::select(date, city_name, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
#   pivot_longer(cols = -c(date, city_name)) %>%
#   dplyr::mutate(name = case_when(
#     name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
#     name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
#   )) %>%
#   dplyr::filter(city_name %in% top_adm) %>%
#   ggplot() +
#   geom_line(aes(x = date,
#                 y = value,
#                 color = name),
#             size = 1) +
#   labs(color = NULL,
#        x = NULL,
#        y = "NTL") +
#   facet_wrap(~city_name) +
#   scale_color_manual(values = c(wbg_color_light,
#                                 wbg_color_dark)) +
#   theme_manual +
#   theme(legend.position = "top")
```

### Monthly

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "bordercrossings", "_", "monthly", ".Rds")))

ntl_df <- ntl_df %>% 
  dplyr::mutate(date = date %>% ymd())

ntl_df %>%
  filter(border_country %in% "Iraq") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Iraq") +
  theme_manual

ntl_df %>%
  filter(border_country %in% "Jordan") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Jordan") +
  theme_manual

ntl_df %>%
  filter(border_country %in% "Lebanon") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Lebanon") +
  theme_manual

ntl_df %>%
  filter(border_country %in% "Turkey") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Turkey") +
  theme_manual
```

### Daily

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "bordercrossings", "_", "daily", ".Rds")))

ntl_df <- ntl_df %>% 
  dplyr::mutate(date = date %>% ymd()) %>%
  dplyr::filter(date >= ymd(DAILY_START))

ntl_df %>%
  filter(border_country %in% "Iraq") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Iraq") +
  theme_manual

ntl_df %>%
  filter(border_country %in% "Jordan") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Jordan") +
  theme_manual

ntl_df %>%
  filter(border_country %in% "Lebanon") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Lebanon") +
  theme_manual

ntl_df %>%
  filter(border_country %in% "Turkey") %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~crossing_name_1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       title = "Border Crossings with Turkey") +
  theme_manual
```

```{r}
# ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
#                             paste0("ntl_", "bordercrossings", "_", "monthly", ".Rds")))
# 
# ntl_df %>%
#   mutate(date = date %>% ymd()) %>%
#   filter(date >= ymd("2018-01-01")) %>%
#   dplyr::rename(city_name = name) %>%
#   dplyr::select(date, city_name, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
#   pivot_longer(cols = -c(date, city_name)) %>%
#   dplyr::mutate(name = case_when(
#     name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
#     name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
#   )) %>%
#   dplyr::filter(city_name %in% top_adm) %>%
#   ggplot() +
#   geom_line(aes(x = date,
#                 y = value,
#                 color = name),
#             size = 1) +
#   labs(color = NULL,
#        x = NULL,
#        y = "NTL") +
#   facet_wrap(~city_name) +
#   scale_color_manual(values = c(wbg_color_light,
#                                 wbg_color_dark)) +
#   theme_manual +
#   theme(legend.position = "top")
```

## Top Cities

### Annual

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "citiesbuff", "_", "annual", ".Rds")))

top_adm <- ntl_df %>%
  filter(date == 2023) %>%
  arrange(-ntl_sum) %>%
  pull(geoname_id) %>%
  head(12)

ntl_df %>%
  filter(geoname_id %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~name) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "cities5km", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::rename(city_name = name) %>%
  dplyr::select(date, city_name, geoname_id, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, city_name, geoname_id)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  dplyr::filter(geoname_id %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            size = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~city_name) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

### Monthly

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "citiesbuff", "_", "monthly", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd("2018-01-01")) %>%
  filter(geoname_id %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~name) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "cities5km", "_", "monthly", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd("2018-01-01")) %>%
  dplyr::rename(city_name = name) %>%
  dplyr::select(date, city_name, geoname_id, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, city_name, geoname_id)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  dplyr::filter(geoname_id %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            size = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~city_name) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

### Daily

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "citiesbuff", "_", "daily", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd(DAILY_START)) %>%
  filter(geoname_id %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~name) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "cities5km", "_", "daily", ".Rds")))

ntl_df %>%
  mutate(date = date %>% ymd()) %>%
  filter(date >= ymd(DAILY_START)) %>%
  dplyr::rename(city_name = name) %>%
  dplyr::select(date, city_name, geoname_id, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, city_name, geoname_id)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  dplyr::filter(geoname_id %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            size = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~city_name) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

# % Change Maps: From 2019

## ADM 1
```{r}

## Load
adm_sf <- load_boundaries(1)
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm1_annual.Rds"))

## Prep data
ntl_wide_df <- ntl_df %>%
  dplyr::filter(date %in% c(pc_base_year:2023)) %>%
  dplyr::select(uid, date, ntl_sum) %>%
  dplyr::mutate(date = paste0("yr", date)) %>%
  pivot_wider(names_from = date,
              values_from = ntl_sum,
              id_cols = uid)

ntl_wide_df$base_year <- ntl_wide_df[[paste0("yr", pc_base_year)]]

ntl_wide_df <- ntl_wide_df %>%
  dplyr::mutate(pc20 = (yr2020 - base_year)/base_year * 100,
                pc21 = (yr2021 - base_year)/base_year * 100,
                pc22 = (yr2022 - base_year)/base_year * 100,
                pc23 = (yr2023 - base_year)/base_year * 100)

## Map
adm_data_sf <- adm_sf %>%
  left_join(ntl_wide_df, by = "uid") %>%
  pivot_longer(cols = c(pc20, pc21, pc22, pc23)) %>%
  dplyr::mutate(name = case_when(
    name == "pc20" ~ paste0("% Change: ",pc_base_year," to 2020"),
    name == "pc21" ~ paste0("% Change: ",pc_base_year," to 2021"),
    name == "pc22" ~ paste0("% Change: ",pc_base_year," to 2022"),
    name == "pc23" ~ paste0("% Change: ",pc_base_year," to 2023")
  ))

adm_data_sf$value[adm_data_sf$value >= 100] <- 100
adm_data_sf$value[adm_data_sf$value <= -100] <- -100

ggplot() +
  geom_sf(data = adm_data_sf,
          aes(fill = value)) +
  facet_wrap(~name) +
  labs(fill = "% Change") +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "forestgreen",
                       midpoint = 0,
                       limits = c(-100, 100),
                       breaks = c(-100, -50, 0, 50, 100),
                       labels = c("< -100", "-50", "0", "50", "> 100")) +
  theme_manual +
  theme_void() +
  theme(strip.text = element_text(face = "bold"))
```

## ADM 2
```{r}

## Load
adm_sf <- load_boundaries(2)
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm2_annual.Rds"))

## Prep data
ntl_wide_df <- ntl_df %>%
  dplyr::filter(date %in% c(pc_base_year:2023)) %>%
  dplyr::select(uid, date, ntl_sum) %>%
  dplyr::mutate(date = paste0("yr", date)) %>%
  pivot_wider(names_from = date,
              values_from = ntl_sum,
              id_cols = uid)

ntl_wide_df$base_year <- ntl_wide_df[[paste0("yr", pc_base_year)]]

ntl_wide_df <- ntl_wide_df %>%
  dplyr::mutate(pc20 = (yr2020 - base_year)/base_year * 100,
                pc21 = (yr2021 - base_year)/base_year * 100,
                pc22 = (yr2022 - base_year)/base_year * 100,
                pc23 = (yr2023 - base_year)/base_year * 100)

## Map
adm_data_sf <- adm_sf %>%
  left_join(ntl_wide_df, by = "uid") %>%
  pivot_longer(cols = c(pc20, pc21, pc22, pc23)) %>%
  dplyr::mutate(name = case_when(
    name == "pc20" ~ paste0("% Change: ",pc_base_year," to 2020"),
    name == "pc21" ~ paste0("% Change: ",pc_base_year," to 2021"),
    name == "pc22" ~ paste0("% Change: ",pc_base_year," to 2022"),
    name == "pc23" ~ paste0("% Change: ",pc_base_year," to 2023")
  ))

adm_data_sf$value[adm_data_sf$value >= 100] <- 100
adm_data_sf$value[adm_data_sf$value <= -100] <- -100

ggplot() +
  geom_sf(data = adm_data_sf,
          aes(fill = value)) +
  facet_wrap(~name) +
  labs(fill = "% Change") +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "forestgreen",
                       midpoint = 0,
                       limits = c(-100, 100),
                       breaks = c(-100, -50, 0, 50, 100),
                       labels = c("< -100", "-50", "0", "50", "> 100")) +
  theme_manual +
  theme_void() +
  theme(strip.text = element_text(face = "bold"))
```

# % Change Maps: 2022 - 2023, 2023 - 2024

## Raster

```{r}
make_change_raster <- function(r_end, r_base, THRESH){
  r_cng <- r_base
  
  r_cng[] <- r_end[] - r_base[]
  
  r_cng[][ (r_cng[] < THRESH) & (r_cng[] > -THRESH) ] <- 0
  r_cng[][r_cng[] >= THRESH]  <- 1
  r_cng[][r_cng[] <= -THRESH] <- -1
  r_cng[][ (r_base[] < THRESH) ] <- NA
  
  r_cng_df <- r_cng %>% 
    rasterToPoints() %>%
    as.data.frame() 
  
  r_cng_df$layer <- r_cng_df[,3]
  
  r_cng_df <- r_cng_df %>%
    dplyr::mutate(layer = case_when(
      layer == 1 ~ "New Lights",
      layer == 0 ~ "No Change",
      layer == -1 ~ "Lights Lost"
    ))
  
  
  return(r_cng_df)
}

roi_sf <- load_gadm(0)

#### 2023 to 2023
r_2023 <- file.path(ntl_bm_dir, "rasters", "monthly") %>%
  list.files(full.names = T) %>%
  str_subset("2023_01|2023_02|2023_03|2023_04|2023_05|2023_06") %>%
  rast() %>%
  mean() %>%
  raster() %>%
  mask(roi_sf)

r_2024 <- file.path(ntl_bm_dir, "rasters", "monthly") %>%
  list.files(full.names = T) %>%
  str_subset("2024_01|2024_02|2024_03|2024_04|2024_05|2024_06") %>%
  rast() %>%
  mean() %>%
  raster() %>%
  mask(roi_sf)

change_24_23 <- make_change_raster(r_2024, r_2023, 1) %>%
  mutate(title = "2023 to 2024")

#### 2022 to 2023
r_2022 <- file.path(ntl_bm_dir, "rasters", "annual") %>%
  list.files(full.names = T) %>%
  str_subset("2022") %>%
  raster() %>%
  mask(roi_sf)

r_2023 <- file.path(ntl_bm_dir, "rasters", "annual") %>%
  list.files(full.names = T) %>%
  str_subset("2023") %>%
  raster() %>%
  mask(roi_sf)

change_23_22 <- make_change_raster(r_2023, r_2022, 1) %>%
  mutate(title = "2022 to 2023")

#### Maps
r_df <- bind_rows(change_24_23,
                  change_23_22)

r_df %>%
  ggplot() +
  geom_sf(data = roi_sf) +
  geom_raster(aes(x = x, y = y,
                  fill = layer)) +
  facet_wrap(~title) +
  labs(fill = NULL) +
  theme_void() +
  theme(strip.text = element_text(face = "bold"),
        legend.position = "bottom") 
```

## ADM 1

```{r}
## Load
adm_sf <- load_boundaries(1)
ntl_monthly_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm1_monthly.Rds"))
ntl_annual_df  <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm1_annual.Rds"))

## Prep data: 23 to 24
ntl_wide_23_24_df <- ntl_monthly_df %>%
  dplyr::mutate(month = date %>% month(),
                year = date %>% year()) %>%
  dplyr::filter(month <= 6,
                year %in% c(2023, 2024)) %>%
  group_by(uid, year) %>%
  dplyr::summarise(ntl_sum = mean(ntl_sum)) %>%
  ungroup() %>%
  pivot_wider(id_cols = uid,
              names_from = year,
              values_from = ntl_sum) %>%
  dplyr::mutate(change = `2024` - `2023`,
                pchange = change / `2023` * 100) %>%
  mutate(title = "2023 to 2024")

adm_data_24_sf <- adm_sf %>%
  left_join(ntl_wide_23_24_df, by = "uid")

## Prep data: 22 to 23
ntl_wide_22_23_df <- ntl_annual_df %>%
  dplyr::filter(date %in% c(2022, 2023)) %>%
  pivot_wider(id_cols = uid,
              names_from = date,
              values_from = ntl_sum) %>%
  dplyr::mutate(change = `2023` - `2022`,
                pchange = change / `2022` * 100) %>%
  mutate(title = "2022 to 2023")

adm_data_23_sf <- adm_sf %>%
  left_join(ntl_wide_22_23_df, by = "uid")

## Append
adm_data_sf <- bind_rows(adm_data_23_sf,
                         adm_data_24_sf)

adm_data_sf$pchange[adm_data_sf$pchange >= 100]  <- 100
adm_data_sf$pchange[adm_data_sf$pchange <= -100] <- -100

adm_data_sf %>%
  ggplot() +
  geom_sf(aes(fill = pchange)) +
  facet_wrap(~title) +
  scale_fill_gradient2(
    low = "red",       # color for minimum (-100)
    mid = "white",     # color for zero (0)
    high = "green",    # color for maximum (100)
    midpoint = 0,      # where white should be (neutral point)
    limits = c(-100, 100)) +
  theme_void() +
  labs(fill = "% Change") +
  theme(strip.text = element_text(face = "bold"),
        legend.position = "bottom")
```

## ADM 2

```{r}
## Load
adm_sf <- load_boundaries(2)
ntl_monthly_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm2_monthly.Rds"))
ntl_annual_df  <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm2_annual.Rds"))

## Prep data: 23 to 24
ntl_wide_23_24_df <- ntl_monthly_df %>%
  dplyr::mutate(month = date %>% month(),
                year = date %>% year()) %>%
  dplyr::filter(month <= 6,
                year %in% c(2023, 2024)) %>%
  group_by(uid, year) %>%
  dplyr::summarise(ntl_sum = mean(ntl_sum)) %>%
  ungroup() %>%
  pivot_wider(id_cols = uid,
              names_from = year,
              values_from = ntl_sum) %>%
  dplyr::mutate(change = `2024` - `2023`,
                pchange = change / `2023` * 100) %>%
  mutate(title = "2023 to 2024")

adm_data_24_sf <- adm_sf %>%
  left_join(ntl_wide_23_24_df, by = "uid")

## Prep data: 22 to 23
ntl_wide_22_23_df <- ntl_annual_df %>%
  dplyr::filter(date %in% c(2022, 2023)) %>%
  pivot_wider(id_cols = uid,
              names_from = date,
              values_from = ntl_sum) %>%
  dplyr::mutate(change = `2023` - `2022`,
                pchange = change / `2022` * 100) %>%
  mutate(title = "2022 to 2023")

adm_data_23_sf <- adm_sf %>%
  left_join(ntl_wide_22_23_df, by = "uid")

## Append
adm_data_sf <- bind_rows(adm_data_23_sf,
                         adm_data_24_sf)

adm_data_sf$pchange[adm_data_sf$pchange >= 100]  <- 100
adm_data_sf$pchange[adm_data_sf$pchange <= -100] <- -100

adm_data_sf %>%
  ggplot() +
  geom_sf(aes(fill = pchange)) +
  facet_wrap(~title) +
  scale_fill_gradient2(
    low = "red",       # color for minimum (-100)
    mid = "white",     # color for zero (0)
    high = "green",    # color for maximum (100)
    midpoint = 0,      # where white should be (neutral point)
    limits = c(-100, 100)) +
  labs(fill = "% Change") +
  theme_void() +
  theme(strip.text = element_text(face = "bold"),
        legend.position = "bottom")
```

## ADM 3

```{r}
## Load
adm_sf <- load_boundaries(3)
ntl_monthly_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm3_monthly.Rds"))
ntl_annual_df  <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm3_annual.Rds"))

## Prep data: 23 to 24
ntl_wide_23_24_df <- ntl_monthly_df %>%
  dplyr::mutate(month = date %>% month(),
                year = date %>% year()) %>%
  dplyr::filter(month <= 6,
                year %in% c(2023, 2024)) %>%
  group_by(uid, year) %>%
  dplyr::summarise(ntl_sum = mean(ntl_sum)) %>%
  ungroup() %>%
  pivot_wider(id_cols = uid,
              names_from = year,
              values_from = ntl_sum) %>%
  dplyr::mutate(change = `2024` - `2023`,
                pchange = change / `2023` * 100) %>%
  mutate(title = "2023 to 2024")

adm_data_24_sf <- adm_sf %>%
  left_join(ntl_wide_23_24_df, by = "uid")

## Prep data: 22 to 23
ntl_wide_22_23_df <- ntl_annual_df %>%
  dplyr::filter(date %in% c(2022, 2023)) %>%
  pivot_wider(id_cols = uid,
              names_from = date,
              values_from = ntl_sum) %>%
  dplyr::mutate(change = `2023` - `2022`,
                pchange = change / `2022` * 100) %>%
  mutate(title = "2022 to 2023")

adm_data_23_sf <- adm_sf %>%
  left_join(ntl_wide_22_23_df, by = "uid")

## Append
adm_data_sf <- bind_rows(adm_data_23_sf,
                         adm_data_24_sf)

adm_data_sf$pchange[adm_data_sf$pchange >= 100]  <- 100
adm_data_sf$pchange[adm_data_sf$pchange <= -100] <- -100

adm_data_sf %>%
  ggplot() +
  geom_sf(aes(fill = pchange)) +
  facet_wrap(~title) +
  scale_fill_gradient2(
    low = "red",       # color for minimum (-100)
    mid = "white",     # color for zero (0)
    high = "green",    # color for maximum (100)
    midpoint = 0,      # where white should be (neutral point)
    limits = c(-100, 100)) +
  labs(fill = "% Change") +
  theme_void() +
  theme(strip.text = element_text(face = "bold"),
        legend.position = "bottom")
```

# Impact of Lebanon Conflict

## ADM 1
```{r}
## Load
lbn_sf <- read_sf(file.path(data_dir, "geoBoundaries-LBN-ADM0-all",
                            "geoBoundaries-LBN-ADM0_simplified.shp"))
adm_sf <- load_boundaries(1)
ntl_monthly_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm1_monthly.Rds"))

adm_sf$dist_lebanon <- adm_sf %>% st_distance(lbn_sf) %>% as.vector()
adm_sf <- adm_sf %>%
  dplyr::filter(dist_lebanon <= 1000)

ntl_monthly_df[ntl_monthly_df$ADM1_EN %in% adm_sf$ADM1_EN,] %>%
  dplyr::mutate(date = date %>% ymd()) %>%
  dplyr::filter(date >= ymd("2024-01-01")) %>%
  ggplot() +
  geom_vline(xintercept = ymd("2024-09-01"), color = "red") +
  geom_col(aes(x = date,
                y = ntl_sum)) +
  labs(x = NULL,
       y = "NTL Radiance") +
  facet_wrap(~ADM1_EN) +
  theme_classic2() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))
```

## ADM 2
```{r}
## Load
lbn_sf <- read_sf(file.path(data_dir, "geoBoundaries-LBN-ADM0-all",
                            "geoBoundaries-LBN-ADM0_simplified.shp"))
adm_sf <- load_boundaries(2)
ntl_monthly_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm2_monthly.Rds"))

adm_sf$dist_lebanon <- adm_sf %>% st_distance(lbn_sf) %>% as.vector()
adm_sf <- adm_sf %>%
  dplyr::filter(dist_lebanon <= 1000)

ntl_monthly_df[ntl_monthly_df$ADM2_EN %in% adm_sf$ADM2_EN,] %>%
  dplyr::mutate(date = date %>% ymd()) %>%
  dplyr::filter(date >= ymd("2024-01-01")) %>%
  ggplot() +
  geom_vline(xintercept = ymd("2024-09-01"), color = "red") +
  geom_col(aes(x = date,
                y = ntl_sum)) +
  labs(x = NULL,
       y = "NTL Radiance") +
  facet_wrap(~ADM2_EN) +
  theme_classic2() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))
```

## ADM 3
```{r}
## Load
lbn_sf <- read_sf(file.path(data_dir, "geoBoundaries-LBN-ADM0-all",
                            "geoBoundaries-LBN-ADM0_simplified.shp"))
adm_sf <- load_boundaries(3)
ntl_monthly_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm3_monthly.Rds"))

adm_sf$dist_lebanon <- adm_sf %>% st_distance(lbn_sf) %>% as.vector()
adm_sf <- adm_sf %>%
  dplyr::filter(dist_lebanon <= 1000)

ntl_monthly_df[ntl_monthly_df$ADM3_EN %in% adm_sf$ADM3_EN,] %>%
  dplyr::mutate(date = date %>% ymd()) %>%
  dplyr::filter(date >= ymd("2024-01-01")) %>%
  ggplot() +
  geom_vline(xintercept = ymd("2024-09-01"), color = "red") +
  geom_col(aes(x = date,
                y = ntl_sum)) +
  labs(x = NULL,
       y = "NTL Radiance") +
  facet_wrap(~ADM3_EN) +
  theme_classic2() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))
```

## ADM 4
```{r}
#| fig-height: 40

## Load
lbn_sf <- read_sf(file.path(data_dir, "geoBoundaries-LBN-ADM0-all",
                            "geoBoundaries-LBN-ADM0_simplified.shp"))
adm_sf <- load_boundaries(4)
ntl_monthly_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm4_5km_monthly.Rds"))

adm_sf$dist_lebanon <- adm_sf %>% st_distance(lbn_sf) %>% as.vector()
adm_sf <- adm_sf %>%
  dplyr::filter(dist_lebanon <= 1000)

ntl_monthly_df[ntl_monthly_df$uid %in% adm_sf$uid,] %>%
  dplyr::mutate(date = date %>% ymd()) %>%
  dplyr::filter(date >= ymd("2024-01-01")) %>%
  ggplot() +
  geom_vline(xintercept = ymd("2024-09-01"), color = "red") +
  geom_col(aes(x = date,
                y = ntl_sum)) +
  labs(x = NULL,
       y = "NTL Radiance") +
  facet_wrap(~admin4Name_en,
             ncol = 5) +
  theme_classic2() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))
```