---
title: "Nighttime Lights Analysis"
author: "Data Lab"
date: "9/20/2024"
format:
  html:
    toc: true
    code-fold: true
knitr:
  opts_chunk:
    warning: false
    message: false
---

Nighttime lights have become a commonly used resource to estimate changes in local economic activity. This document analyzes nighttime lights within the country.

# Data

We use nighttime lights data from [VIIRS Black Marble](https://blackmarble.gsfc.nasa.gov/). Raw nighttime lights data requires correction due to cloud cover and stray light, such as lunar light. The Black Marble dataset applies [advanced algorithms](https://www.sciencedirect.com/science/article/pii/S003442571830110X) to correct raw nighttime light values and calibrate data so that trends in lights over time can be meaningfully analyzed. From VIIRS Black Marble, we use data from January 2012 through present—where data is available at a 500-meter resolution.

# Methodology

Within different units of analysis (e.g, administrative units) we use the sum of nighttime lights. Where relevant, we distinguish between lights generated from gas flaring and lights removing gas flaring. We use the [World Bank's Global Gas Flaring Tracker](https://www.worldbank.org/en/programs/gasflaringreduction/global-flaring-data) which indicates the location of gas flaring locations. When removing gas flaring lights, we remove lights within 10km of a gas flaring location; when looking at lights in gas flaring locations, we take the sum of lights within 10km of gas flaring locations. 

# Code Setup

```{r}
#| warning: false
#| message: false

#### MAIN PARAMS
proj_dir     <- "/Users/rmarty/Library/CloudStorage/OneDrive-SharedLibraries-WBG/Development Data Partnership - Syria Economic Monitor"
iso3_code    <- "SYR"
iso2_code    <- "SY"
pc_base_year <- 2019

#### Packages
library(tidyverse)
library(sf)
library(leaflet)
library(leaflet.providers)
library(ggpubr)
library(terra)
library(tidyterra)
library(gtools)
library(readxl)
library(janitor)
library(geodata)
library(exactextractr)
library(blackmarbler)
library(dplyr)
library(readxl)
library(janitor)
library(ggpubr)
library(WDI)
library(kableExtra)
library(elevatr)
library(ggnewscale)
library(pals)
library(haven)

#### Paths

## Define
data_dir       <- file.path(proj_dir, "Data")
ntl_bm_dir     <- file.path(data_dir, "Nighttime Lights BlackMarble")
gas_flare_dir  <- file.path(data_dir, "gas-flaring")
bound_dir      <- file.path(data_dir, "boundaries")
unocha_dir <- file.path(data_dir, "UNOCHA")

#### Theme
theme_manual <- theme_classic2() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))

#### Load GADM Functions
load_gadm <- function(i){
  
  if(i == 0){
    gadm_sf <- read_sf(file.path(unocha_dir, "syr_pplp_adm4_unocha_20210113", 
                                 "syr_admbnda_adm0_uncs_unocha_20201217.json"))
    gadm_sf$date <- NULL
  }
  
  if(i == 1){
    gadm_sf <- read_sf(file.path(unocha_dir, "syr_pplp_adm4_unocha_20210113", 
                                 "syr_admbnda_adm1_uncs_unocha_20201217.json")) 
    gadm_sf$date <- NULL
  }
  
  if(i == 2){
    gadm_sf <- read_sf(file.path(unocha_dir, "syr_pplp_adm4_unocha_20210113", 
                                 "syr_admbnda_adm2_uncs_unocha_20201217.json")) 
    gadm_sf$date <- NULL
  }
  
  if(i == 3){
    gadm_sf <- read_sf(file.path(unocha_dir, "syr_pplp_adm4_unocha_20210113",
                                 "syr_admbnda_adm3_uncs_unocha_20201217.json")) 
    gadm_sf$date <- NULL
  }
  
  gadm_sf
}

#### Colors
wbg_color_light <- "#1AA1DB"
wbg_color_dark <- "#02224B"
```

# Maps of Nighttime Lights

## Interactive Map

```{r}

#| warning: false
#| message: false

## Load boundaries
adm0_sf <- load_gadm(0)

## Load/prep raster
prep_r <- function(year_i){
  r <- rast(file.path(ntl_bm_dir, "rasters", "annual",
                      paste0("VNP46A4_NearNadir_Composite_Snow_Free_qflag_t",year_i,".tif")))
  r <- r %>% mask(adm0_sf)
  r[][r[] <= 0.05] <- NA
  r[] <- log(r[] + 1)
  #r[] <- log(r[] + 1)
  return(r)
}

r_2012 <- prep_r(2012)
r_2013 <- prep_r(2013)
r_2014 <- prep_r(2014)
r_2015 <- prep_r(2015)
r_2016 <- prep_r(2016)
r_2017 <- prep_r(2017)
r_2018 <- prep_r(2018)
r_2019 <- prep_r(2019)
r_2020 <- prep_r(2020)
r_2021 <- prep_r(2021)
r_2022 <- prep_r(2022)
r_2023 <- prep_r(2023)
r_2024 <- prep_r(2024)

## Make map
pal <- colorNumeric(c("black", "yellow", "red"), unique(c(r_2012[],
                                                          r_2013[],
                                                          r_2014[],
                                                          r_2015[],
                                                          r_2016[],
                                                          r_2017[],
                                                          r_2018[],
                                                          r_2019[],
                                                          r_2020[],
                                                          r_2021[],
                                                          r_2022[],
                                                          r_2023[],
                                                          r_2024[])),
                    na.color = "transparent")

leaflet() %>%
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addRasterImage(r_2012, colors = pal, opacity = 1, group = "2012") %>%
  addRasterImage(r_2013, colors = pal, opacity = 1, group = "2013") %>%
  addRasterImage(r_2014, colors = pal, opacity = 1, group = "2014") %>%
  addRasterImage(r_2015, colors = pal, opacity = 1, group = "2015") %>%
  addRasterImage(r_2016, colors = pal, opacity = 1, group = "2016") %>%
  addRasterImage(r_2017, colors = pal, opacity = 1, group = "2017") %>%
  addRasterImage(r_2018, colors = pal, opacity = 1, group = "2018") %>%
  addRasterImage(r_2019, colors = pal, opacity = 1, group = "2019") %>%
  addRasterImage(r_2020, colors = pal, opacity = 1, group = "2020") %>%
  addRasterImage(r_2021, colors = pal, opacity = 1, group = "2021") %>%
  addRasterImage(r_2022, colors = pal, opacity = 1, group = "2022") %>%
  addRasterImage(r_2023, colors = pal, opacity = 1, group = "2023") %>%
  addRasterImage(r_2024, colors = pal, opacity = 1, group = "2024") %>%
  addLayersControl(
    baseGroups = paste0(2012:2024),
    options = layersControlOptions(collapsed=FALSE)
  )
```


## Static Map of Nighttime Lights

```{r}
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

## Load boundaries
adm0_sf <- load_gadm(0)

## Elevation
elev_r <- rast(file.path(data_dir, "elevation", "elev.tiff"))
slope_r <- terrain(elev_r)
slope_r <- slope_r %>% mask(adm0_sf)

## Latest Year
r <- rast(file.path(ntl_bm_dir, "rasters", "annual",
                    paste0("VNP46A4_NearNadir_Composite_Snow_Free_qflag_t",2024,".tif")))
r <- r %>% mask(adm0_sf)

r[][r[] <= 1] <- NA
r[] <- log(r[] + 1)
r[] <- log(r[] + 1)

ocean_colors <- ocean.deep(100)
ocean_colors[98:100] <- "#FFFFFF"

ggplot() +
  geom_spatraster(data = slope_r, 
                  show.legend = FALSE,
                  maxcell = 9e+07) +
  scale_fill_gradient(low = "#49232E",
                      high = "black",
                      na.value = "transparent") +
  ggnewscale::new_scale_fill() +
  geom_spatraster(data = r,
                  aes(fill = t2024),
                  maxcell = 9e+07) +
  scale_fill_gradientn(colors = ocean_colors,
                       #midpoint = 2,
                       na.value = "transparent") +
  labs(fill = "Light Intensity",
       title = "Nighttime Lights",
       subtitle = 2024) +
  coord_sf() +
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
        plot.subtitle = element_text(face = "italic", hjust = 0.5, size = 16),
        legend.position = c(0.7, 0.15)) +
  guides(fill = guide_colorbar(direction = "horizontal",
                               label = FALSE,         
                               title.position = "top"))
```
# Diagnostics

## Nighttime Lights as Proxy for GDP
```{r}
# Download GDP
dir.create(file.path(data_dir, "wdi"))
OUT_FILE <- file.path(data_dir, "wdi", "gdp.Rds")

if(!file.exists(OUT_FILE)){
  gdp_df <- WDI(indicator = c("NY.GDP.PCAP.CD",
                              "NY.GDP.MKTP.CD"), 
                country = iso3_code, 
                start = 2012, 
                end = 2024)
  
  saveRDS(gdp_df, OUT_FILE)
} else{
  gdp_df <- readRDS(OUT_FILE)
}

# Correlation with GDP
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("adm0", "_", "annual", ".Rds")))

ntl_df <- ntl_df %>%
  dplyr::rename(year = date)

# Filter
if(paste(iso3_code, collapse = ";") %in% "SSD;SDN"){
  ntl_df <- ntl_df[ntl_df$COUNTRY %in% "Sudan",]
  gdp_df <- gdp_df[gdp_df$country %in% "Sudan",]
}

ntl_gdp_df <- ntl_df %>%
  left_join(gdp_df, by = "year") %>%
  clean_names()

## Line Plots
ntl_gdp_df %>%
  dplyr::select(year, 
                viirs_bm_sum,
                ny_gdp_pcap_cd,
                ny_gdp_mktp_cd) %>%
  pivot_longer(cols = -year) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_sum" ~ "Nighttime Lights",
    name == "ntl_gf_10km_sum" ~ "Nighttime Lights:\nGas Flares",
    name == "ntl_nogf_10km_sum" ~ "Nighttime Lights:\nNo Gas Flares",
    name == "ny_gdp_pcap_cd" ~ "GDP, Per Capita",
    name == "ny_gdp_mktp_cd" ~ "GDP"
  )) %>%
  dplyr::mutate(type = name %>% str_detect("GDP")) %>%
  ggplot(aes(x = year,
             y = value,
             color = type)) +
  geom_line(linewidth = 1) +
  facet_wrap(~name,
             scales = "free_y") +
  labs(x = NULL,
       y = NULL) +
  scale_color_manual(values = c(wbg_color_light, wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "none")

## Correlation: GDP
ntl_gdp_df %>%
  dplyr::select(viirs_bm_sum, 
                ny_gdp_mktp_cd) %>%
  pivot_longer(cols = -c(ny_gdp_mktp_cd)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_sum" ~ "NTL"
  )) %>%
  
  ggplot(aes(x = value,
             y = ny_gdp_mktp_cd)) +
  geom_point() +
  geom_smooth(method=lm, 
              se=FALSE,
              color = wbg_color_light) +
  stat_cor() +
  labs(x = "NTL",
       y = "GDP",
       title = "Correlation of Nighttime Lights with GDP") +
theme_manual 

## Correlation: GDP
ntl_gdp_df %>%
  dplyr::select(viirs_bm_sum, 
                ny_gdp_pcap_cd) %>%
  pivot_longer(cols = -c(ny_gdp_pcap_cd)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_sum" ~ "NTL"
  )) %>%
  
  ggplot(aes(x = value,
             y = ny_gdp_pcap_cd)) +
  geom_point() +
  geom_smooth(method=lm, 
              se=FALSE,
              color = wbg_color_light) +
  stat_cor() +
  labs(x = "NTL",
       y = "GDP, per Capita",
       title = "Correlation of Nighttime Lights with per Capita GDP") +
  theme_manual
```


# Trends in Nighttime Lights

## Annual vs Monthly
```{r}
# day_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
#                             paste0("ntl_", "adm0", "_", "daily", ".Rds"))) %>%
#   dplyr::mutate(date = date %>% ymd())

month_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                              paste0("adm0", "_", "monthly", ".Rds"))) %>%
  dplyr::mutate(date = date %>% ymd())

year_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                             paste0("adm0", "_", "annual", ".Rds")))

year_df$date <- paste0(year_df$date, "-01-01") %>% ymd()

ggplot() +
  geom_col(data = year_df,
           aes(x = date, y = viirs_bm_sum, color = "Annual"),
           fill = "gray90") +
  # geom_line(data = day_df,
  #           aes(x = date, y = ntl_sum, color = "Daily"),
  #           linewidth = 0.1) +
  geom_line(data = month_df,
            aes(x = date, y = viirs_bm_sum, color = "Monthly"),
            linewidth = 1) +
  labs(x = NULL,
       y = "Nighttime Lights",
       color = NULL) +
  theme_manual
```

## Annual

### ADM0

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm0_annual.Rds"))

ntl_df %>%
  ggplot(aes(x = date,
             y = viirs_bm_sum)) +
  geom_col(fill = wbg_color_dark) +
  labs(x = NULL,
       y = "Radiance") +
  theme_manual
```

### ADM1

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm1_annual.Rds"))

ntl_df %>%
  ggplot(aes(x = date,
             y = viirs_bm_sum)) +
  geom_col(fill = wbg_color_dark) +
  labs(x = NULL,
       y = "Radiance") +
  theme_manual +
  facet_wrap(~adm1_en)
```

### ADM2

```{r}
#| fig-width: 8
#| fig-height: 10

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm2_annual.Rds"))

ntl_df %>%
ggplot(aes(x = date,
y = viirs_bm_sum)) +
geom_col(fill = wbg_color_dark) +
labs(x = NULL,
y = "Radiance") +
theme_manual +
facet_wrap(~adm2_en)
```

### ADM3

```{r}
#| fig-width: 8
#| fig-height: 20

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm3_annual.Rds"))

ntl_df %>%
ggplot(aes(x = date,
y = viirs_bm_sum)) +
geom_col(fill = wbg_color_dark) +
labs(x = NULL,
y = "Radiance") +
theme_manual +
facet_wrap(~adm3_en)
```

### Border Crossings

```{r}
#| fig-width: 8
#| fig-height: 12

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "border_crossings_1km_annual.Rds"))

ntl_df %>%
  ggplot(aes(x = date,
  y = viirs_bm_sum)) +
  geom_col(fill = wbg_color_dark) +
  labs(x = NULL,
  y = "Radiance") +
  theme_manual +
  facet_wrap(~crossing, ncol = 4)
```

## Monthly

### ADM0

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm0_monthly.Rds"))

ntl_df %>%
ggplot(aes(x = date,
y = viirs_bm_sum)) +
geom_col(fill = wbg_color_dark) +
labs(x = NULL,
y = "Radiance") +
theme_manual
```

### ADM1

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm1_monthly.Rds"))

ntl_df %>%
ggplot(aes(x = date,
y = viirs_bm_sum)) +
geom_col(fill = wbg_color_dark) +
labs(x = NULL,
y = "Radiance") +
theme_manual +
facet_wrap(~adm1_en)
```

### ADM2

```{r}
#| fig-width: 8
#| fig-height: 10

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm2_monthly.Rds"))

ntl_df %>%
ggplot(aes(x = date,
y = viirs_bm_sum)) +
geom_col(fill = wbg_color_dark) +
labs(x = NULL,
y = "Radiance") +
theme_manual +
facet_wrap(~adm2_en)
```

### ADM3

```{r}
#| fig-width: 8
#| fig-height: 20

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm3_monthly.Rds"))

ntl_df %>%
ggplot(aes(x = date,
y = viirs_bm_sum)) +
geom_col(fill = wbg_color_dark) +
labs(x = NULL,
y = "Radiance") +
theme_manual +
facet_wrap(~adm3_en)
```

### Border Crossings

```{r}
#| fig-width: 8
#| fig-height: 12

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "border_crossings_1km_monthly.Rds"))

ntl_df %>%
  ggplot(aes(x = date,
  y = viirs_bm_sum)) +
  geom_col(fill = wbg_color_dark) +
  labs(x = NULL,
  y = "Radiance") +
  theme_manual +
  facet_wrap(~crossing, ncol = 4)
```

## Daily

### ADM0

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm0_daily.Rds"))

ntl_df %>%
ggplot(aes(x = date,
y = viirs_bm_sum)) +
geom_col(fill = wbg_color_dark) +
labs(x = NULL,
y = "Radiance") +
theme_manual
```

### ADM1

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm1_daily.Rds"))

ntl_df %>%
ggplot(aes(x = date,
y = viirs_bm_sum)) +
geom_col(fill = wbg_color_dark) +
labs(x = NULL,
y = "Radiance") +
theme_manual +
facet_wrap(~adm1_en)
```

### ADM2

```{r}
#| fig-width: 8
#| fig-height: 10

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm2_daily.Rds"))

ntl_df %>%
ggplot(aes(x = date,
y = viirs_bm_sum)) +
geom_col(fill = wbg_color_dark) +
labs(x = NULL,
y = "Radiance") +
theme_manual +
facet_wrap(~adm2_en)
```

### ADM3

```{r}
#| fig-width: 8
#| fig-height: 20

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm3_daily.Rds"))

ntl_df %>%
ggplot(aes(x = date,
y = viirs_bm_sum)) +
geom_col(fill = wbg_color_dark) +
labs(x = NULL,
y = "Radiance") +
theme_manual +
facet_wrap(~adm3_en)
```

### Border Crossings

```{r}
#| fig-width: 8
#| fig-height: 12

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "border_crossings_1km_daily.Rds"))

ntl_df %>%
  ggplot(aes(x = date,
  y = viirs_bm_sum)) +
  geom_col(fill = wbg_color_dark) +
  labs(x = NULL,
  y = "Radiance") +
  theme_manual +
  facet_wrap(~crossing, ncol = 4)
```

# % Change Maps

## ADM 1
```{r}

## Load
adm_sf <- load_gadm(1) %>% clean_names()
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm1_annual.Rds"))

## Prep data
ntl_wide_df <- ntl_df %>%
  dplyr::filter(date %in% c(pc_base_year:2024)) %>%
  dplyr::select(adm1_en, date, viirs_bm_sum) %>%
  dplyr::mutate(date = paste0("yr", date)) %>%
  pivot_wider(names_from = date,
              values_from = viirs_bm_sum,
              id_cols = adm1_en)

ntl_wide_df$base_year <- ntl_wide_df[[paste0("yr", pc_base_year)]]

ntl_wide_df <- ntl_wide_df %>%
  dplyr::mutate(pc20 = (yr2020 - base_year)/base_year * 100,
                pc21 = (yr2021 - base_year)/base_year * 100,
                pc22 = (yr2022 - base_year)/base_year * 100,
                pc23 = (yr2023 - base_year)/base_year * 100,
                pc24 = (yr2024 - base_year)/base_year * 100)

## Map
adm_data_sf <- adm_sf %>%
  left_join(ntl_wide_df, by = "adm1_en") %>%
  pivot_longer(cols = c(pc21, pc22, pc23, pc24)) %>%
  dplyr::mutate(name = case_when(
    name == "pc21" ~ paste0("% Change: ",pc_base_year," to 2021"),
    name == "pc22" ~ paste0("% Change: ",pc_base_year," to 2022"),
    name == "pc23" ~ paste0("% Change: ",pc_base_year," to 2023"),
    name == "pc24" ~ paste0("% Change: ",pc_base_year," to 2024")
  ))

adm_data_sf$value[adm_data_sf$value >= 100] <- 100
adm_data_sf$value[adm_data_sf$value <= -100] <- -100

ggplot() +
  geom_sf(data = adm_data_sf,
          aes(fill = value)) +
  facet_wrap(~name) +
  labs(fill = "% Change") +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "forestgreen",
                       midpoint = 0,
                       limits = c(-100, 100),
                       breaks = c(-100, -50, 0, 50, 100),
                       labels = c("< -100", "-50", "0", "50", "> 100")) +
  theme_manual +
  theme_void() +
  theme(strip.text = element_text(face = "bold"))
```
## ADM 2
```{r}

## Load
adm_sf <- load_gadm(2) %>% clean_names()
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "adm2_annual.Rds"))

## Prep data
ntl_wide_df <- ntl_df %>%
  dplyr::filter(date %in% c(pc_base_year:2024)) %>%
  dplyr::select(adm2_en, date, viirs_bm_sum) %>%
  dplyr::mutate(date = paste0("yr", date)) %>%
  pivot_wider(names_from = date,
              values_from = viirs_bm_sum,
              id_cols = adm2_en)

ntl_wide_df$base_year <- ntl_wide_df[[paste0("yr", pc_base_year)]]

ntl_wide_df <- ntl_wide_df %>%
  dplyr::mutate(pc20 = (yr2020 - base_year)/base_year * 100,
                pc21 = (yr2021 - base_year)/base_year * 100,
                pc22 = (yr2022 - base_year)/base_year * 100,
                pc23 = (yr2023 - base_year)/base_year * 100,
                pc24 = (yr2024 - base_year)/base_year * 100)

## Map
adm_data_sf <- adm_sf %>%
  left_join(ntl_wide_df, by = "adm2_en") %>%
  pivot_longer(cols = c(pc21, pc22, pc23, pc24)) %>%
  dplyr::mutate(name = case_when(
    name == "pc21" ~ paste0("% Change: ",pc_base_year," to 2021"),
    name == "pc22" ~ paste0("% Change: ",pc_base_year," to 2022"),
    name == "pc23" ~ paste0("% Change: ",pc_base_year," to 2023"),
    name == "pc24" ~ paste0("% Change: ",pc_base_year," to 2024")
  ))

adm_data_sf$value[adm_data_sf$value >= 100] <- 100
adm_data_sf$value[adm_data_sf$value <= -100] <- -100

ggplot() +
  geom_sf(data = adm_data_sf,
          aes(fill = value)) +
  facet_wrap(~name) +
  labs(fill = "% Change") +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "forestgreen",
                       midpoint = 0,
                       limits = c(-100, 100),
                       breaks = c(-100, -50, 0, 50, 100),
                       labels = c("< -100", "-50", "0", "50", "> 100")) +
  theme_manual +
  theme_void() +
  theme(strip.text = element_text(face = "bold"))
```